\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{alltt}
\usepackage{fancyvrb}
\usepackage[svgnames,xcdraw,table]{xcolor}
\usepackage{multirow}
\usepackage{xnewcommand}
\usepackage{longtable}
\usepackage[font=small,skip=1pt]{caption}
\usepackage{subcaption, balance, appendix}
\usepackage{geometry, setspace, fullpage}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage{alltt}


\begin{document}

\newcommand[\global]{\rcow}{\rowcolors[\hline]{1}{gray!10}{Snow}\arrayrulecolor{Black}\setlength\arrayrulewidth{1pt}}

\title{CS423 : MP2 Documentation}
\author{{\bf Group 2}: Debjit Pal (dpal2), Anirudh Jayakumar (ajayaku2), Neha Chaube (nchaub2) \\ and Divya Balakrishna (dbalakr2)}
\date{\today}

\maketitle

\section{Introduction}
We describe the important data structures and the functionalities of the functions we created to build up {\em Rate Monotonic Scheduler} (RMS). We used a multi-file approach to keep the modularity of the design and to make sure that all the group members can work in parallel without harming others' code. Hence we changed the {\tt Makefile} provided in MP1. In Section~\ref{subsec:Files}, we enlist the different files relevant to this work, in Section~\ref{subsec:Make}, we describe the changes that we made in the {\tt Makefile}. In Section~\ref{subsec:pcb} we demonstrate the augmented process control block (PCB) that we created for this project. Section~\ref{subsec:Proc}, we demonstrate the creation of {\tt proc} file system, the APIs to {\tt read} from and {\tt write} into the {\tt proc} filesystem. In this section we also demonstrate the usage of the {\tt procfile\_write} call back for process {\em registration, de-registration} and {\em yield}. In Section~\ref{subsec:admission} we demonstrate the {\tt admission\_control} procedure which in turn used by the {\tt procfile\_write} function to check if a new process can be admitted without violating deadlines of any already registered process. In this section we also describe {\tt remove\_task} which removes the task from the linked list when it is done with its computation. In Section~\ref{subsec:Link} we describe the functions that are needed to handle linklists in the Kernel. Section~\ref{subsec:dispatch} demonstrates the working of the dispatching thread and enlists the necessary functions we wrote. Section~\ref{sec:toRun} demonstrates step-by-step running procedure of the RMS module and the application. Section~\ref{sec:Conclude} concludes the documents along with our learnings and the steps we took to debug the code.

%% This part needs to be written after I receive Aniruddhs, Neha, Divya's writing 

\subsection{Files included in the MP2 Project}\label{subsec:Files}

The list of the included files for this MP is given in Table~\ref{table:filelist}.

\begin{table*}[h]
  \centering
  \rcow
  \caption{File list included in the project\label{table:filelist}}
  \begin{tabular}{|p{4cm}|p{8cm}|}
    {\tt Makefile}  &  To build up the Kernel module and user application called {\em my\_factorial}  \\
    {\tt structure.h} & Defining the augmented process control block, the functions and the other data types and data structures relevant to this project\\
    {\tt mp2.c}     &  Includes the kernel init and exit modules, proc file creation, proc file read and write function, admission control  \\
    {\tt linklist.h} & Declares all interface functions to read and write into the linklist\\
    {\tt linklist.c} & Definitions of all the interface functions to read and write in the linked list \\
    {\tt thread.c}  & Implements the dispatcher thread \\
    {\tt thread.h}  & Contains the function declarations for the dispatcher thread implementation \\
    {\tt process.c} & Implementation of the parametrized factorial user application \\
  \end{tabular}
\end{table*}

\subsection{Important Changes in Makefile}\label{subsec:Make}

We took a multi-file approach to make sure each of the group member can work on different parts of the implementation, we modified the given {\tt Makefile} from MP1 in the following way:

\begin{verbatim}
 obj-m += mp2_final.o
 mp2_final-objs := linklist.o thread.o mp2.o
\end{verbatim}

The name of the object module has been changed to {\tt mp2\_final.o} as mentioned in the {\tt obj-m}. That {\tt mp2\_final.o} contains three different object modules as mentioned in the {\tt mp2\_final-objs}. The {\em ordering} of the object module is important here because mp2.o contains a call to a dispatch thread function that is defined in thread.o. Hence, thread.c needs to be compiled before else gcc will give a compiler error message {\tt
"implicit function declaration found"}. Further, mp2.c contains the {\tt MODULE\_LICENSE} declaration and hence it needs to be compiled at the very end after all other C files are compiled else gcc will pop an error message {\tt MODULE\_LICENSE declaration not found}. If we load such a kernel module without {\tt MODULE\_LICENSE} definition, the kernel will get {\em tainted} and may crash. In our project, the name of the kernel module is {\tt mp2\_final.ko} which we load in the kernel using {\em insmod} command.

\nopagebreak
\input{Debjit}
\input{Anirudh}

%\include{Neha}
\input{Divya}

\section{User Application}

We implemented an user application which is called {\tt process} and built whenever the kernel module is built using the {\tt Makefile}.  The application takes three parameters namely
\begin{enumerate}
	\item Period
	\item Computation time
	\item Number of jobs
\end{enumerate}
A typical invocation of the process looks like:

\[\underbrace{./process}~~\underbrace{1000}_{\text{period}}~~\underbrace{100}_{\text{computation}} ~~\underbrace{100}_{\text{number of jobs}}\]

The process implements the following functions as shown in Table~\ref{table:process}.

\begin{table*}[h]
	\centering
	\rcow
	\caption{List of functions implemented in the user application named {\tt process}\label{table:process}}
	\begin{tabular}{|p{7.5cm}|p{8cm}|}
		{\tt void myregister(pid\_t pid, unsigned long int period, unsigned long int computation, char *file);}  &  Registers the process in the proc file system with the user specified parameters.  \\
		{\tt int read\_status(char *file);}     & Read the status of the registration of the process from proc file system. If successfully registered returns size of the proc size buffer read else returns {\tt 0}  \\
		{\tt unsigned long factorial(unsigned long i);}&  Calculates factorial of a pre-specified number\\
		{\tt void yield(pid\_t pid,char *file);}&  Sends request for yield\\
		{\tt void unregister(pid\_t pid, char *file) ;}  & When the process is complete with user specified number of jobs, this functions request for de-registration through the proc file system \\
	\end{tabular}
\end{table*}

\section{How to run the program}\label{sec:toRun}

Please use the following steps to compile, insert and remove the kernel module and to run our application {\em my\_factorial } program.

\begin{enumerate}
	\item To compile the kernel module and the user application:
	\begin{verbatim}
	# make
	\end{verbatim}
	This will create {\tt mp2\_final.ko} kernel object module and the user application {\em my\_factorial}.
	\item To insert the kernel module:
	\begin{verbatim}
	# sudo insmod mp2_final.ko
	\end{verbatim}
	This should print a few confirmation messages in the {\bf /var/log/kern.log} file confirming that the kernel module has been loaded successfully.
	\item To run the user application:
	\begin{verbatim}
	# ./process 10 500 100 (./process <period> <computation> <number of jobs>)
	\end{verbatim}
	\item To output the CPU time of the running processes in konsole:
	\begin{verbatim}
	# cat /proc/mp2/status
	\end{verbatim}
	\item To unload the kernel module:
	\begin{verbatim}
	# sudo rmmod mp2_final
	\end{verbatim}
	This will print a few confirmation messages in the {\bf /var/log/kern.log} file confirming that the kernel module has been unloaded successfully.
\end{enumerate}


\section{Conclusion}\label{sec:Conclude}

In this MP we implement a kernel module performing real time scheduling following the principle of Rate Monotonic Scheduler. We also implement a test application to test the kernel module. In this MP we learnt
\begin{enumerate}
	\item To use the real time priorities
	\item Use of dispatching thread for context switch implementation, to send the process from the READY to RUNNING and SLEEPING states.
	\item We learnt the usage of the timer.
\end{enumerate}


\end{document} 